---
title: "Meta-analysis using metafor"
author: "Vernon Visser"
date: "21 June 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to meta-analysis

In this Markdown document I provide a little bit of code to help get you started with doing a basic meta-anlysis.

The example used here is from [Gouda-Vossos et al. (2018)](https://doi.org/10.1007/s40750-018-0099-y) and deals with mate choice in humans. The
authors did a meta-analysis of experiments in which the "attractiveness" of a person was rated before and after a treatment, which involved either "addition" or "augmentation" designs. In this example we are only going to look at their experiment (i), which was an "addition" experiment in which the attractiveness of females was rated when they were alone (the "control") and then when they were surrounded by a number of men (the "treatment").

![An example of a "mate choice experiment" in humans. From [Dunn & Doria (2010)](http://dx.doi.org/10.1037/h0099305).](Mate_choice_addition.png)
  
In this Toolbox RMarkdown document, you will find out how to:

  * Calculate effect sizes
  * Run fixed-effect and random-effects meta-analytic models
  * Run a meta-regression
  * Produce forest plots
  * Check for publication bias


## Getting started in R

Load packages and get data

```{r, message=F}
library(metafor) #Install this package first if you do not have it
dat = read.csv('Gouda-Vossos_S2.csv')
head(dat)
```

## Calculate effect sizes

We use the log of the response ratio (lnRR), which is the mean of the treatment (M_T) divided by the mean of the control (M_C)

```{r}
dat = escalc(measure="ROM", m1i=Mean_with, m2i=Mean_without, sd1i=SD_with, sd2i=SD_without, n1i=Number_with,
           n2i=Number_without, data=dat)
head(dat)
```

## Fixed-effect and random-effects meta-analytic models for (i) 

In the paper, the authors present the results of a random-effects meta-analytic model, which is more conservative and accounts for 
non-independence of effect sizes (i.e. assumes the studies do not come from the same population, which is a realistic assumption).

The fixed-effect model is merely provided for comparison.

```{r}
#Random-effects model
model_i_re = rma(yi, vi, data=dat, method='REML')
summary(model_i_re)

#Fixed-effect model
model_i_fe = rma(yi, vi, data=dat, method='FE')
summary(model_i_fe)
```

## Meta-regression model for (i)

Moderators are what are termed predictors in ,for example, multiple regression. In the Gouda-Vossos et al. (2018) paper they did not use any 
moderators because their sample sizes were too low, but below is an example of what a meta-regression would look like with "No_of_Stim" as
a predictor.

```{r}
model_i_mod = rma(yi, vi, mods = No_of_Stim, data=dat, method='REML')
summary(model_i_mod)
```


## Forest plot

Forest plots are a standard way of presenting meta-analysis results. The individual study effect sizes are shown with their corresponding
95% confidence intervals, as well as the overall effect size at the bottom of the plot.

```{r}
forest(model_i_re, slab = dat$First_Author, pch=16)
```

## Publication bias

It is important to test for publication bias in meta-analyses. There are numerous sources of publication bias, but can include, for example, the tendency to only report studies with significant results or large effects, or the difficulty in acquiring studies in the grey or foreign literature.

Funnel plots are one of the most commonly used methods to look for publication bias. Essentially, the funnel plot provides an indication of whether there are "gaps" in the studies used to conduct the meta-analysis. Very often studies with low sample sizes (y-axis) and small effect sizes (x-axis) are missing.

```{r}
#Funnel plot
funnel(model_i_re)
```

Trim-and-fill analyses calculate the number of "missing" studies. They do this by removing the smallest studies (trimming) and recalculating
the overall effect size (filling) until there is symmetry. A number of estimators of the number of missing studies are used, termed L, R & Q. Below, you can see the results for all three of these estimators.

```{r}
#Trim and fill
tfL0 = trimfill(model_i_re, estimator="L0")
tfL0

tfR0 = trimfill(model_i_re, estimator="R0")
tfR0

tfQ0 = trimfill(model_i_re, estimator="Q0")
tfQ0
```

## Further reading

This brief introduction obviously ignores many important concepts and methods available for meta-analysis. For more information on this
subject, I suggest reading:

  * [Nakagawa et al. 2017. BMC Ecology 15:18](https://doi.org/10.1186/s12915-017-0357-7)
  * [Nakagawa & Cuthill 2007. Biological Reviews 82](https://doi.org/10.1111/j.1469-185X.2007.00027.x)
  * [Harrison 2011. Methods in Ecology and Evolution 2](https://doi.org/10.1111/j.2041-210X.2010.00056.x)
  * [Koricheva, Gurevitch & Mengersen 2013. Handbook of Meta-analysis in Ecology and Evolution](https://press.princeton.edu/titles/10045.html)
  * [Moher et al. 2009. Annals of Internal Medicine 151](https://doi.org/10.7326/0003-4819-151-4-200908180-00135)
